---
title: "Homework 2"
author: Yuanming Mao
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df =
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)


precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1,
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df =
  bind_rows(precip_2018, precip_2017)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. he dataset contains information on year, month, and trash collected, include some specific kinds of trash. 

The **trashwheel_df** dataset has `r nrow(trashwheel_df)` observations. It includes date of trash collection for each dumpster, weight and volume of trash collected, and the number of specific types of trash.

The **precip_df** dataset includes month precipitation data, and has `r nrow(precip_df)` observations.

For available data, the total precipitation in 2018 was `r sum(pull(precip_2018, var = total))` inches. The median number of sports balls in a dumpster in 2017 was `r filter(trashwheel_df, year == 2017) %>% pull(sports_balls) %>% median()`.


## problem 2

Read NYC Transit data

```{r}
transit_df = 
  read.csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:ada, -exit_only, -staffing, -staff_hours) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0)) %>% 
  mutate(entry = as.logical(entry))

```

**Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?**


There are `r distinct(transit_df, line, station_name, .keep_all = TRUE) %>% nrow()` distinct stations.

`r distinct(transit_df, line, station_name, .keep_all = TRUE) %>% filter(ada == TRUE) %>% nrow()` stations are ADA compliant.

`r nrow(filter(transit_df, vending == "NO", entry == TRUE))/nrow(filter(transit_df, vending == "NO"))*100`% of station entrances / exits without vending allow entrance.


Reformat data so that route number and route name are distinct variables. 

```{r}
transit_tidy_data = 
  transit_df %>% 
  mutate_at(vars(route1:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  )
```

`r distinct(transit_tidy_data, line, station_name, .keep_all = TRUE) %>% filter(route_name == "A") %>% nrow()` distinct stations serve the A train.

`r distinct(transit_tidy_data, line, station_name, .keep_all = TRUE) %>% filter(route_name == "A", ada == TRUE) %>% nrow()` are ADA compliant of the stations that serve the A train.



